---
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
library(MASS)
library(ResourceSelection) # For Hosmer-Lemeshow test
library(ggplot2)
library(tidyverse)
library(bestglm)
library(pROC)
library(corrplot)
library(ggplot2)
library(ggcorrplot)
library(dplyr)
library(corrr)
library(glmnet)
library(mlbench)
library(haven)
library(survival)
library(survminer)
```

### Load Data Set

Load in the "House Prices - Advanced Regression Techniques" data set
from kaggle.

```{r Load Data}
# training data set
iowa.homes <- read.csv("data/train.csv", header=TRUE, sep=",") 

```

### Explore Missing Values Part 1

Below we can see that there are a number of variables with missing data.
The most concerning features are Alley, LotFrontage, FireplaceQu, Fence,
MiscFeature, as they have very large amounts (15% - 99%) of data absent.
For Alley, MiscFeature, PoolQC, FireplaceQu, Fence, these are homes that
do not have an alley, a miscelaneous feature (ie. elevator, 2nd Garage,
Shed, Tennis Court, other), a pool, fireplace, or a fence. LotFrontage
represents the amount of space between the street and the start of the
lot, and its possible that these NA values are due to no measurement
being made or no space between the lot and the street. The features with
40% or more missing data (Alley, MiscFeature, PoolQC, FireplaceQu, and
Fence) will be removed and new binary features (HasAlley,
HasMiscFeature, HasFireplace, HasFence) will be added.

```{r, missing value information}
# get percentage of missing data for each column
na.percentages <- sort(colMeans(is.na(iowa.homes)) * 100)

#create df of missing data percentage information
na.df <- data.frame(
  feature = names(na.percentages),
  percentage_na = na.percentages
)

#subset df to only contain features with some missing data
na.df <- na.df[na.df$percentage_na > 0,]

#plot feature by missing data percentage to visualize
ggplot(data=na.df, aes(x = reorder(feature, percentage_na), y = percentage_na)) + 
  geom_col(fill = "lightblue") +
  theme_dark() +
  theme(axis.text.x=element_text(angle=90), plot.title = element_text(hjust = 0.5)) +
  labs(x = "Feature Name", y = "% Missing", title = "Features With Missing Data")

#print data frame to display numeric percentages for each feature
na.df
```

### Feature Engineer Part 1

Below I will be adding new binary features HasAlley, HasMiscFeature,
HasFireplace, HasFence, HasPool, HasBasement and removing those with
large amounts of missing data (NA \>= 40%).

```{r, Feature Engineer Part 1}
features_to_remove <- na.df$feature[na.df$percentage_na > 40]

func_new_binary_vars <- function(df){
  #add new binary feature "HasBsmt"
  df <- df %>%
    mutate(HasAlley = case_when(
      is.na(Alley) ~ 'no', .default = 'yes')
    )
  
  #add new binary feature "HasMiscFeature"
  df <- df %>%
    mutate(HasMiscFeature = case_when(
      is.na(MiscFeature) ~ 'no', .default = 'yes')
    )
  
  #add new binary feature "HasFireplace"
  df <- df %>%
    mutate(HasFireplace = case_when(
      is.na(FireplaceQu) ~ 'no', .default = 'yes')
    )
  
  #add new binary feature "HasFence"
  df <- df %>%
    mutate(HasFence = case_when(
      is.na(Fence) ~ 'no', .default = 'yes')
    )
  
  #add new binary feature "HasPool"
  df <- df %>%
    mutate(HasPool = case_when(
      is.na(PoolQC) ~ 'no', .default = 'yes')
    )
  
  #add new binary feature "HasBsmt"
  df <- df %>%
    mutate(HasBsmt = case_when(
      is.na(BsmtQual) ~ 'no', .default = 'yes')
    )
  
  #remove columns with large amounts of missing data
  df <- df[, colMeans(is.na(df)) <= 0.4]

  return(df)
}

iowa.homes <- func_new_binary_vars(iowa.homes)
```

### Visualize New Binary Features And Other Categorical Features with Missing Data

From the visualizations below, we can see that there is not a huge
difference in means for a house having a miscellaneous feature or not,
having an alley or not, having a fence or not. However, there seems to
be a significant difference in means for a house that has a pool vs.
not, where homes with a pool seems to be significantly more expensive,
and homes with a fireplace are also more expensive. I will therefore
remove the HasAlley, HasFence, and HasMiscFeature columns, but will keep
the HasBsmt, HasPool, and HasFireplace features.

I also decided to look at the other two binary vars, 'Street',
'Utilities', and 'CentralAir' to visualize any difference in the
saleprice between the two levels. Because there are so few values in the
NoSeWa and because the mean sale price of this variable appears to be
fairly close to the mean sale price of AllPub, I will remove this
feature from the data set. However, the difference in mean and variance
from analysis of the box plots for 'Street' and 'CentralAir' do seem to
be statistically significant and indicate that these features may be
associated with home sale price.

```{r, Visualize New Binary Features}

# create list of variables to plot
bin_cols_to_plot = c("HasMiscFeature", "HasAlley", "HasFence", "HasFireplace", "HasPool", "HasBsmt", "Street", "Utilities", "CentralAir")

# turn new binary variables to factor type
iowa.homes[bin_cols_to_plot] <- lapply(iowa.homes[bin_cols_to_plot], as.factor)

# create box plots for binary variables of interest
lapply(bin_cols_to_plot, 
      function(n)
        ggplot(data=iowa.homes, aes_string(x=n, y='SalePrice')) + geom_boxplot()
      )

func_remove_features_pt2 <- function(df){
  df <- subset(df, select = -c(HasAlley, HasFence, HasMiscFeature, Street, Utilities))
  return(df)
}

iowa.homes <- func_remove_features_pt2(iowa.homes)
```

```{r, get numeric and categorical columns}

numeric_cols <- iowa.homes %>% select(where(is.numeric), -SalePrice) %>% names()
categorical_cols <- iowa.homes %>% select(where(~ is.factor(.) || is.character(.))) %>% names()
# categorical_cols <- sapply(iowa.homes, negate(is.numeric))

```

### Replace Categorical NAs with 'None'

Below I replace NA with "None" in remaining features that represent the
feature not being present so we do not remove homes where this data is
missing.

```{r Replace categorical NAs with Mode}
func_replace_miss_data <- function(df){
  to_replace <- c("GarageType","GarageFinish","GarageQual","GarageCond",
                 "BsmtFinType1","BsmtFinType2","BsmtExposure","BsmtCond","BsmtQual",
                 "MasVnrType")
  
  # replace missing data with 'None' in specified columns
  df <- df %>% 
    mutate_at(to_replace, ~replace_na(.,"None"))
  
  return(df) 
  }

iowa.homes <- func_replace_miss_data(iowa.homes)
  
```

### Remaining Features With Missing Data

As we can see, we have managed to take care of a large number of the
variables that had missing data, especially those with large percents of
missing data. However, we still have to decide how to treat missing data
in 'Electrical', 'MassVnrArea', 'GarageYearBuilt', and 'LotFrontage'.

```{r missing data}
# get percentage of missing data for each column
na.percentages <- sort(colMeans(is.na(iowa.homes)) * 100)

#create df of missing data percentage information
na.df <- data.frame(
  feature = names(na.percentages),
  percentage_na = na.percentages
)

#subset df to only contain features with some missing data
na.df <- na.df[na.df$percentage_na > 0,]
na.df

#plot feature by missing data percentage to visualize
ggplot(data=na.df, aes(x = reorder(feature, percentage_na), y = percentage_na)) + 
  geom_col(fill = "lightblue") +
  theme_dark() +
  theme(axis.text.x=element_text(angle=90), plot.title = element_text(hjust = 0.5)) +
  labs(x = "Feature Name", y = "% Missing", title = "Features With Missing Data")

```

### Replace NA's in Remaining Categorical Features and Some Numerical Features

Below I replace NA's in the categorical features with NA's in the
training data set listed above. I also went ahead and replaced missing
data for features that have missing data in the validate and test set in
order to clean these data sets.

I replace many of the categorical features with the mode of the feature.
I also replace NA in some of the numerical features with 0, (as a 0
represents the feature not being present such as GarageArea, if this is
NA, it means the home does not have a garage and therefore its area can
be represented with 0).

```{r replace NAs in GarageYrBlt}

# create function to get mode of categorical column of interest
func_mode<-function(x){
 
  z<-data.frame(Var=x)%>%
    group_by(Var)%>%
    summarise(Total_Count=n())%>%
    ungroup()%>%
    head(1)%>%
    select(Var)%>%
    pull(Var)
  
  return(z)
}

func_replace_miss_data_pt2 <- function(df){
  # set to 0 representing homes without a garage
  df <- df %>%
    mutate(GarageYrBlt=replace_na(GarageYrBlt, 0))

  df <- df %>%
    mutate(GarageArea=replace_na(GarageArea, 0))
  
  df <- df %>%
    mutate(Exterior1st=replace_na(Exterior1st, func_mode(Exterior1st)))

  df <- df %>%
    mutate(Exterior2nd=replace_na(Exterior2nd, func_mode(Exterior2nd)))
  
  df <- df %>%
    mutate(MSZoning=replace_na(MSZoning, func_mode(MSZoning)))

  df <- df %>%
    mutate(KitchenQual=replace_na(KitchenQual, func_mode(KitchenQual)))
    
  df <- df %>%
    mutate(GarageArea=replace_na(GarageArea, 0))
  # replace missing data in 'Electrical' feature with "no_info"
  df <- df %>%
    mutate(Electrical=replace_na(Electrical, "no_info"))
  
  #replace missing data in 'MasVnrArea' with 0, as this is not part of home
  df <- df %>%
    mutate(MasVnrArea=replace_na(MasVnrArea, 0))
  
  df <- df %>%
    mutate(TotalBsmtSF = replace_na(TotalBsmtSF, 0))
  
  #replace missing 'LotFrontage' data with median
  df <- df %>%
    mutate(LotFrontage=replace_na(LotFrontage, median(df$LotFrontage, na.rm = TRUE)))
  
  return(df)
}

iowa.homes <- func_replace_miss_data_pt2(iowa.homes)
```

### Check for Missing Data

Now we can see that there are no longer any columns with missing data in
the training data.

```{r}
# get percentage of missing data for each column
na.percentages <- sort(colMeans(is.na(iowa.homes)) * 100)

#create df of missing data percentage information
na.df <- data.frame(
  feature = names(na.percentages),
  percentage_na = na.percentages
)

#subset df to only contain features with some missing data
na.df <- na.df[na.df$percentage_na > 0,]
na.df
```

### Set Categorical Variables to Factor Type

Below I create a function to set the categorical features to factor type
for modeling later on.

```{r set cat features to factor type vars}
func_cat_to_factor<-function(df){
  categorical_cols <- df %>% select(where(~ is.factor(.) || is.character(.))) %>% names()
  df[categorical_cols] <- lapply(df[categorical_cols], as.factor)
  
  return(df)
}

iowa.homes<-func_cat_to_factor(iowa.homes)
```

### Create New Numeric Features (TotalSF, HouseAge, RemodelAge)

Below I have created a new feature "TotalSF" which represents the total
surface area of each home. I felt that this would be a good feature to
include in the correlation analysis in the next step. I also create a
feature called 'HouseAge' representing the age of the house, and
'RemodelAge' representing the age of the remodel if one was done.

```{r}
# sum relevant square footage measurements
func_feature_eng <- function(df){
  df <- df %>%
    mutate(TotalSF = rowSums(df[, c("TotalBsmtSF", "X1stFlrSF", "X2ndFlrSF", "X3SsnPorch")]))
  
  df <- df%>%
    mutate(HouseAge = YrSold - YearBuilt)
  
  df <- df %>% 
    mutate(RemodelAge = YrSold - YearRemodAdd)
  
  return(df)
}

iowa.homes <- func_feature_eng(iowa.homes)
```

### Numeric Feature Correlation With SalePrice

Below I have run a correlation analysis on the numeric features with
sale price. All features with correlations greater than 0.5 will be
included in the step-wise model selection portion of this project. Given
that 'HouseAge' shows a stronger correlation than 'YrSold' or
'YearBuilt' on their own, I will use 'HouseAge' in their place.
Similarly, 'RemodelAge' shows a tighter correlation with 'SalePrice'
than 'YearRemodAdd' alone, and I will use 'RemodelAge' in its place.
Finally, 'TotalSF' shows a stronger correlation with 'TotalBsmtSF',
'X1stFlrSF', X2ndFlrSF', and 'X3SsnPorch' on their own and I will use
'TotalSF' in their place.

```{r Numeric Feature Corr with SalePrice}

#create correlation matrix that focuses on variables correlation to SalePrice
correlation_matrix <- iowa.homes %>% select(where(is.numeric)) %>% correlate() %>% focus(SalePrice)

#plot correlations on bar plot
correlation_matrix %>%
  ggplot(aes(x = term, y = SalePrice)) +
    geom_bar(stat = "identity") +
    ylab("Correlation with Sale Price") +
    xlab("Variable") +
    theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))

#get list of features with correlation > 0.5 for numeric features
correlated_numeric_cols <- c(correlation_matrix$term[abs(correlation_matrix$SalePrice) >= 0.5]) 
correlated_numeric_cols

```

### Correlated Variables

Below I look at the correlation between GarageArea and GarageCars to see
if one can be used in place of the other. Indeed, they have a
correlation of 0.88. I will use GarageArea in place of GarageCars.

```{r look at correaltion between variables}
cor(iowa.homes$GarageArea, iowa.homes$GarageCars)
```

### Get List of Unnecessary Numeric Features

Below, using the correlation matrix focused on SalePrice, I get a list
of numerical features that are less correlated (corr \< 0.5) with
SalePrice to remove these features at a later step.

```{r clean numeric features from df}

correlation_matrix <- iowa.homes %>% select(where(is.numeric)) %>% correlate() %>% focus(SalePrice)

#get list of features with correlation > 0.5 for numeric features
un_correlated_numeric_cols <- c(correlation_matrix$term[abs(correlation_matrix$SalePrice) < 0.5]) 

un_correlated_numeric_cols
```

### Analysis of Variance Between Mutli-Level Factors

#### Get all Muti-Level Factors

Below I wrote some code to get all the categorical features with more
than 2 levels to run an anova on these features in the next step.

```{r}

res <- sapply(iowa.homes, nlevels)

res.df <- data.frame(
  feature = names(res),
  num_levels = res
)

# subset df to only contain features with more than 2 levels
res.df1 <- (res.df[res.df$num_levels > 2,])
# subset df to only contain binary features
res.df2 <- (res.df[res.df$num_levels == 2,])

# create list of numeric features with more than 2 levels
ml_categorical_cols <- res.df1$feature
# create list of binary numeric features
binary_categorical_cols <- res.df2$feature

```

#### ANOVA on all Multi-Level Categorical Variables

Below I run and ANOVA (without interaction) on all of the categorical
features with more than 2 levels (non-binary). I did this to try and
assess which categorical features show an association with sale price
based on an f-statistic with a p-value \> 0.01. All non-statistically
significant features from this step (based on an alpha of 0.01) will be
removed from the df.

The qqplot indicates the residuals are normally distributed and the
residuals vs. fitted plot shows the residuals have mostly equal variance
with some outliers.

```{r}

# create formula for anova to run (w/o) interaction on the multilevel 
# categorical columns
formula_str <- paste("SalePrice ~ ", paste(ml_categorical_cols[-1], collapse = "+"))
formula_obj <- as.formula(formula_str)

# run anova normally to view summary
cat.aov <- aov(formula_obj, data = iowa.homes)

# run with anova tidy to get a dataframe like object for parsing easier
cat.aov.df <- broom::tidy(aov(formula_obj, data = iowa.homes))
cat.aov.df <- drop_na(cat.aov.df)

# show output of anova
summary(cat.aov)

#creat list of statistically significant categorical columns
un_signif_ml_categrorical_cols <- cat.aov.df$term[cat.aov.df$p.value > 0.001]

#check plots for assumptions
plot(cat.aov)

# show significant multi-level categorical features
un_signif_ml_categrorical_cols

```

### Remove less associated and correlated columns

Below I created a list of all the insignificant numerical features and
non-associated categorical features to be removed from the data frame, I
then remove all of these features from the data frame.

```{r}
func_remove_features_pt3<-function(df){
  to_drop <- c("LandSlope", "BsmtCond", "BsmtFinType2", "Heating", "Electrical", "Functional", "GarageFinish", "GarageQual", "GarageCond","GarageCars", "PavedDrive", "SaleType", "SaleCondition", "MSSubClass", "LotFrontage", "LotArea", "OverallCond", "MasVnrArea", "BsmtFinSF1", "BsmtFinSF2", "BsmtUnfSF", "LowQualFinSF", "BsmtFullBath", "BsmtHalfBath", "HalfBath", "BedroomAbvGr", "KitchenAbvGr", "Fireplaces", "GarageYrBlt", "WoodDeckSF", "OpenPorchSF", "EnclosedPorch", "ScreenPorch", "PoolArea", "MiscVal", "MoSold", "Condition2", "HouseStyle", "RoofMatl", "Exterior1st", "Exterior2nd")
  
  df <- df %>% select(-any_of(to_drop))
  
  return(df)
}

iowa.homes <- func_remove_features_pt3(iowa.homes)

```

### Model Selection

Below I perform a backward model selection, cross validated lasso, and
cross validated ridge regression model selection. I took both the lambda
= min, and lambda = 1se models for ridge and lasso. I then apply these
models to predict values on the validate data set to decide which model
to use on the test data set. I made the decision for the 'best' model
based on the lowest root mean squared error.

### Create full model and intercept model

```{r forward seleciton pt1}

full.model_str <- paste("SalePrice ~ .") 
full.model <- as.formula(full.model_str)

full.model.new <- lm(SalePrice ~ ., data = iowa.homes)
intercept.model <- lm(SalePrice ~ 1, data = iowa.homes)

```

### Backward Subset Selection

```{r}

backward.model <-step(full.model.new,scope=intercept.model,direction="backward",trace=1, data=iowa.homes)
```

### Best Backward Model

Based on the lowest AIC score from the backward model selection, the
best model is the one that included 21 of the 46 variables. Below I
create a linear model and explore its features. Based on the residuals
vs. fitted plot, we can see that the variance is mostly homogenously
spread but gets larger as the fitted values increase. We can also see
that the residuals are mostly normally distributed aside from a tail of
points towards the larger theoretical quantiles.

```{r}
backward.best <- lm(SalePrice ~ LotShape + LandContour + LotConfig + Neighborhood + 
    Condition1 + BldgType + OverallQual + YearBuilt + YearRemodAdd + 
    ExterQual + BsmtQual + BsmtExposure + BsmtFinType1 + TotalBsmtSF + 
    CentralAir + GrLivArea + FullBath + KitchenQual + GarageArea + 
    HasFireplace, data = iowa.homes)

summary(backward.best)
plot(backward.best)
```

### Comparing to Simpler Model

The backward model selection kept the "Condition1" feature. However,
from the model summary above, we can see that none of the levels
(represented by dummy variables) related to this feature are
statistically significant (P\>0.05 for all level of Condition1).
Therefore, below I will test to see if the simpler model without the
'Condition1' feature may actually be better by comparing using an anova.

The anova model comparison gave a p-value of 0.025 which is less than
the alpha of 0.05 and we can conclude that the more complicated model
that includes 'Condition1' should be used.

```{r}
backward.simpler <- lm(SalePrice ~ LotShape + LandContour + LotConfig + Neighborhood + BldgType + OverallQual + YearBuilt + YearRemodAdd + 
    ExterQual + BsmtQual + BsmtExposure + BsmtFinType1 + TotalBsmtSF + 
    CentralAir + GrLivArea + FullBath + KitchenQual + GarageArea + 
    HasFireplace, data = iowa.homes)

anova(backward.best,backward.simpler)
```

### Prepare training data for lasso and ridge regression

```{r prepare data for regression}

X.train = model.matrix(full.model,data=iowa.homes) # Saves the implied model design matrix, which includes dummy coding categorical predictors

X.train<-X.train[,-1] # Remove the first column, which is a column of ones.

y.train = iowa.homes$SalePrice # Saves the outcome variable as a vector


```

### Cross validated lasso selection

We can see from the cross validated lasso selection that we get a lambda
min of 546 and a lambda 1se of 3511. The model generated from the lambda
min removed far less variables than the lambda 1se model. This is
expected because as lambda increases, the penalty against parameter
coefficients is larger and more coefficients are reduced to 0 as lambda
gets larger.

```{r lasso}

set.seed(202211)

cvfit.lasso.train <- cv.glmnet(x=X.train, y=y.train,alpha=1)

plot(cvfit.lasso.train)
```

### cross validated lasso lambda min

```{r lasso min}

# Display your lambda.min here
cvfit.lasso.train$lambda.min


# Display the coefficients associated with lambda.min here
coef(cvfit.lasso.train, s = "lambda.min")
```

### cross validated lasso lambda 1se

```{r lasso 1se}

# Display your lambda.1se here
cvfit.lasso.train$lambda.1se

# Display the coefficients associated with lambda.1se here
coef(cvfit.lasso.train, s = "lambda.1se")

```

### Cross Validated Ridge Regression

Below I perform the cross validated ridge regression. Here we can see
that the lambda min for the ridge regression was 6281 and the lambda 1se
was 123310. Many of the variables are pushed closer to zero in the
lambda 1se model. Again, this is because lambda determines the penalty
size against parameter coefficients and the larger the lambda, the
harsher the penalty. The difference between the ridge regression and the
lasso regression is in the penalty term, where the ridge regression
penalty term pushes coefficients closer to zero as lambda increases, the
lasso regression penalty term can get rid of coefficients all together.

```{r ridge}
set.seed(123456) # Sets a recoverable random seed so ensure that the same results will be obtained if the code is run again

cvfit.ridge.train= cv.glmnet(x=X.train, y=y.train,alpha=0) # Conduct cross-validated ridge regression

plot(cvfit.ridge.train)

```

### cross validated ridge lambda min

```{r}

cvfit.ridge.train$lambda.min # Display the lambda.min value

coef(cvfit.ridge.train, s = "lambda.min") # Display the "shrunk" coefficients associated with lambda.min

summary(cvfit.ridge.train)
```

### cross validated ridge lambda 1se

```{r}
cvfit.ridge.train$lambda.1se # Display the lambda.1se value

coef(cvfit.ridge.train, s = "lambda.1se") # Display the "shrunk" coefficients associated with lambda.1se
```

### Setting up test and validate data

Below I load in the test-validate data set and run the necessary
pre-processing functions on this data set.

```{r}
dat.test.valid <- read.csv("data/test.csv", header=TRUE, sep=",")
dat.saleprice <- read.csv("data/sample_submission.csv")

dat.test.valid <- merge(dat.test.valid, dat.saleprice)

# call necessary functions on dat.test 
dat.test.valid <- func_new_binary_vars(dat.test.valid)

bin_cols_to_plot = c("HasMiscFeature", "HasAlley", "HasFence", "HasFireplace", "HasPool", "HasBsmt", "Street", "Utilities", "CentralAir")

# turn new binary variables to factor type
dat.test.valid[bin_cols_to_plot] <- lapply(dat.test.valid[bin_cols_to_plot], as.factor)

dat.test.valid <- func_remove_features_pt2(dat.test.valid)
dat.test.valid <- func_replace_miss_data(dat.test.valid)
dat.test.valid <- func_replace_miss_data_pt2(dat.test.valid)
dat.test.valid <- func_cat_to_factor(dat.test.valid)
dat.test.valid <- func_feature_eng(dat.test.valid)
dat.test.valid <- func_remove_features_pt3(dat.test.valid)

```

### split data into validate and test 50/50

Below I split the test and validate data 50:50

```{r split data into valid and test}
n <- nrow(dat.test.valid) 

set.seed(202211)
tv.split <- rep(0:1,c(round(n*.5),n-round(n*.5)))
table(tv.split)   

dat.test <- dat.test.valid[tv.split==1,] 
dat.valid <- dat.test.valid[tv.split==0,] 
```

### Prepare validate data for regression analysis

```{r prepare valid data}

X.valid <- model.matrix(full.model,data=dat.valid) # Saves the implied model design matrix, which includes dummy coding categorical predictors

X.valid<-X.valid[,-1] # Remove the first column, which is a column of ones.

# head(X.test) # Confirm that all variables are present and in the expected form - everything looks good!

y.valid = dat.valid$SalePrice # Save the outcome variable as a vector
```

### Running models on validate data

```{r}

backward.best.preds <- predict(backward.best, newx=X.valid)

cvlasso.lambda.min.preds <- predict(cvfit.lasso.train, newx=X.valid, s = cvfit.lasso.train$lambda.min)

cvlasso.lambda.1se.preds <- predict(cvfit.lasso.train, newx=X.valid, s = cvfit.lasso.train$lambda.1se)

cvridge.lambda.min.preds <- predict(cvfit.ridge.train, newx=X.valid, s = cvfit.ridge.train$lambda.min)

cvridge.lambda.1se.preds <- predict(cvfit.ridge.train, newx=X.valid, s = cvfit.ridge.train$lambda.1se)


```

### Comparing root mean squared error from models

Below I compare the root mean square error between the different models
that I have created. We can see that the ridge 1se cross validated model
has the lowest mean squared error term (58,818) and this is the model
that I will select based on this value. The 58,818 means that the model
estimations of sale price in the validation data set, were off by
\$58,818 on average.

Its possible that the ridge model fits best because of the large amount
of features in the data set and there is likely some correlation between
more of the variables than I was able to identify on my own earlier.
This is an example of where the ridge model can be useful as it reduces
coefficients close to zero, but does not get rid of them as with lasso
regression. The lambda 1se being better than the lambda min is likely
due to the large amount of variance and low bias in the training model,
a larger lambda is likely better for predictions as it introduces more
bias while decreasing variance to regularize the model.

```{r model mspe comparison}

backward.best.valid <- sqrt(mean((backward.best.preds-y.valid)^2))
backward.best.valid

rmspe.cvlasso.min.valid <-  sqrt(mean((cvlasso.lambda.min.preds-y.valid)^2)) 
rmspe.cvlasso.min.valid

rmspe.cvlasso.1se.valid <-  sqrt(mean((cvlasso.lambda.1se.preds-y.valid)^2)) 
rmspe.cvlasso.1se.valid

rmspe.cvridge.min.valid <-  sqrt(mean((cvridge.lambda.min.preds-y.valid)^2)) 
rmspe.cvridge.min.valid

rmspe.cvridge.1se.valid <-  sqrt(mean((cvridge.lambda.1se.preds-y.valid)^2)) 
rmspe.cvridge.1se.valid

```

### Testing the Model

Below I first prepare the testing data for predicting and then I use the
selected cross validated ridge lambda 1se model to make SalePrice
predictions using the test data. Finally I get the mean squared error of
the predictions (\$60,141.44), and then I create histograms of the
predicted sale price overlayed by a histogram of the actual sale price
from the test data set.

```{r prepare test data}

X.test <- model.matrix(full.model,data=dat.test) # Saves the implied model design matrix, which includes dummy coding categorical predictors

X.test<-X.test[,-1] # Remove the first column, which is a column of ones.

y.test = dat.test$SalePrice # Save the outcome variable as a vector

```

```{r run model on test}
cvridge.lambda.1se.preds.test <- predict(cvfit.ridge.train, newx=X.test, s = cvfit.ridge.train$lambda.1se)

rmspe.cvridge.1se.test <-  sqrt(mean((cvridge.lambda.1se.preds.test-y.test)^2)) 
rmspe.cvridge.1se.test

```

```{r}
dat.test['PredSalePrice'] <- cvridge.lambda.1se.preds.test
ggplot(data = dat.test) +
 geom_histogram(aes(x = SalePrice, fill = "Actual Sale Price"), alpha = 0.5) +
 geom_histogram(aes(x = PredSalePrice, fill = "Predicted Sale Price"), alpha = 0.5) 

mean(dat.test$SalePrice)
mean(dat.test$PredSalePrice)

mean(dat.test$SalePrice) - mean(dat.test$PredSalePrice)
```

### Conclusion

The root mean squared error of the model predictions of home sale price
using the test data set was 60141.44. This means that the average
difference between the model prediction and the actual sale price was
\$60,141. We can also see the distribution of the predicted sale price
and the actual sale price in the test data in the visual above. Clearly,
the shape of the distributions are noticably different. The predicted
sale price distribution has a relatively large right tail compared with
the actual sale price distribution and the actual sale price
distribution has a larger portion of its data centered around its mean.
While the average difference between the predicted values and the actual
values was 60,141, the mean of the predicated sale price was relatively
close to the mean of the actual sale price. The actual mean was
\$180,775 and the predicted mean was \$175,490, only off by \$5,284.

### Discussion

While this model is not perfect, I still believe it could be a useful
tool for getting a 'ballpark' estimate of home prices in Ames, Iowa. In
the future, I would have liked to find a way feature engineer and deal
with NA values in a more consolidated, streamlined manner. However, I
have much to learn about feature engineering and handling missing data,
and I look forward to becoming better at this. I also would have liked
to have looked at some transformations (perhaps a box-cox
transformation) of the feature of interest (SalePrice) to explore how a
model with such a transformation would compare to the original model.
Finally, looking for interaction between terms would be another useful
step to consider in the future to get a more accurate model.
